Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-06-30 manuscript/pnas_style/manuscript.tex
incorporating extremely helpful reviewer edits. thanks!

@@

\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
roles of genetic drift and selection in shaping genome-wide allele frequency
changes.  Selection can affect allele frequencies, both directly and
indirectly, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome, e.g. linked selection
\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
This approach has shown linked selection has a major role in shaping patterns
of genome-wide diversity across the genomes of a range of sexual species
\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over short
time intervals and its full impact on genome-wide allele frequency changes.

There are numerous examples of rapid phenotypic adaptation
\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact of selection on genome-wide variation
over short timescales in sexual populations remarkably difficult
\cite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
effect of selection on a polygenic trait (such as fitness) is distributed
across numerous loci. This can lead to subtle allele frequency shifts on
standing variation that are difficult to distinguish from background levels of
genetic drift and sampling variance. Increasingly, genomic experimental
evolution studies with multiple timepoints, and in some cases multiple
replicate populations, are being used to detect large-effect selected loci
\cite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
\cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from temporal genomic data is the
temporal autocovariance (i.e. covariance between two timepoints) in allele
frequency changes \cite{Buffalo2019-io}.  These covariances are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are expected to be zero for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a time interval
into the variances and covariances in allele frequency change between
generations.  Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an
artificial selection experiment \cite{Castro2019-uk}. These sequencing
selection experiments have started to uncover selected loci contributing to the
adaptive response; however it is as yet far from clear how much of the allele
frequency genome-wide is driven by selection or genetic drift. We repeatedly
find a signal of temporal covariance, consistent with linked selection acting
to significantly perturb genome-wide allele frequency changes across the genome
in a manner that other approaches would not be able differentiate from genetic
drift. We estimate a lower bound of the fraction of variance in allele
frequency change caused by selection, as well as the correlation between allele
frequency changes between replicate populations caused by convergent selection
pressures. Overall, we demonstrate that linked selection has a powerful role in
shaping genome-wide allele frequency changes over very short timescales.

@@

We first analyzed Barghi et al. (2019), an evolve-and-resequence study with ten
replicate populations exposed to a high temperature lab environment and evolved
for 60 generations, and sequenced every ten generations. Using the seven
timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) of some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalized the entries for
heterozygosity (see SI Appendix, sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
covariances are expected to be zero when only drift is acting, as only
heritable variation for fitness can create covariance between allele frequency
changes in a closed population \cite{Buffalo2019-io}.  Averaging across the ten
replicate temporal covariances matrices, we find temporal covariances that are
statistically significant (95\% block bootstraps CIs do not contain zero),
consistent with linked selection perturbing genome-wide allele frequency
changes over very short time periods. The covariances between all adjacent time
intervals are positive and then decay towards zero as we look at more distant
time intervals (Figure \ref{fig:figure-1} A), as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and the associated additive genetic variance for fitness
decays (which could occur as a population reaches a new optimum)
\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
this general pattern holds; see SI Appendix, Fig. 
\ref{suppfig:barghi-cov-panels}.

Since our covariances are averages over loci, the covariance estimate could be
strongly affected by a few outlier regions. To test whether large outlier
regions drive the genome-wide signal we see in the Barghi et al. (2019)
data, we calculate the covariances in 100kb windows along the genome (we refer
to these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates
(SI Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
covariance, confirming that genomic temporal covariances are non-zero due to
the impact of selection acting across many genomic regions.

@@

  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
    populations, through time from the Barghi et al. (2019) study. Each line
    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
    reference generation $s$ to a later time $t$ which varies along the x-axis;
    each line corresponds to a row of the upper-triangle of the temporal
    covariance matrix with the same color (upper right). The ranges around each
    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
    proportion of the total variance in allele frequency change explained by
    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
    The black line is the $G(t)$ averaged across replicates, with the $95\%$
    block-bootstrap confidence interval. The other lines are the $G(t)$ for
    each individual replicate, with colors indicating what subset of the
    temporal-covariance matrix to the right is being included in the
  calculation of $G(t)$.}

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change: the
ratio of total covariance in allele frequency change to the total variance in
allele frequency change. We denote the change in allele frequency as $\Delta
p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
Since the total variation in allele frequency change can be partitioned into
variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
  \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
\Delta p_j)$ (we bias-correct these for sequencing depth), and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the impact of selection on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time. When the sum of the covariances is
positive, this measure can intuitively be understood as the relative fraction
of allele frequency change normally thought of as ``drift" that is actually due
to selection. Additionally, $G(t)$ can be understood as a short-timescale
estimate of the reduction in neutral diversity due to linked selection (or
equivalently the reduction in neutral effective population size needed to
account linked selection, see SI Appendix, section
\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
sequenced every ten generations, the numerator uses the covariances
  estimated between ten-generation blocks of allele frequency change; thus
  the strong, unobservable, covariances between adjacent generations do not
contribute to the numerator of $G(t)$. Had these covariances been measurable
on shorter timescales, their cumulative effect would likely have been higher
yet (see SI Appendix, sections \ref{supp:barghi-covs} and
\ref{supp:tempblocks} for more details).  Additionally, selection also inflates
the variance in allele frequency change per generation; however, this effect
cannot be easily distinguished from drift. For both these reasons, our measure
$G(t)$ is quite conservative (we demonstrate this through
simulations in SI Appendix, section \ref{supp:tempblocks}).  Still,
we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
allele frequency change over 60 generations is the result of selection (Figure
\ref{fig:figure-1} B). This proportion of variance attributable to selection
builds over time in Figure \ref{fig:figure-1}B as the effects of linked
selection are compounded over the generations unlike genetic drift. Our G(t)
starts to plateau to a constant level as the covariances from earlier
generations have decayed and so no longer contribute as strongly (Figure
\ref{fig:figure-1}).

@@

The replicate design of Barghi et al. (2019) allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are
expected to have allele frequency changes in the same direction if the
selection pressures are similar. Intuitively, where temporal covariances
reflect that alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2}A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2}A). We find these convergent correlation coefficients
are relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while a
substantial fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result consistent with the primary finding
of the original Barghi et al. (2019) study: that alternative loci
contribute to longer term adaptation across the different replicates.

@@

  \caption{{\bf A}: The convergence correlations, averaged across
    Barghi et al. (2019) replicate pairs, through time. Each line
    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
    p_{s})$ from a starting reference generation $s$ to a later time $t$, which
    varies along the x-axis; each line corresponds to a row of the temporal
    convergence correlation matrix depicted to the right (where the
      diagonal elements represent the convergence correlations between the same
      timepoints across replicate populations). We note that convergent
      correlation for the last timepoint is an outlier; we are unsure as to the
      cause of this, e.g. it does not appear to be driven by a single pair of
      replicates. {\bf B}: The convergence correlations between individual
      pairs of replicates in the Kelly and Hughes (2019) data (note the
      confidence intervals are plotted, but are small on this y-axis scale).
      {\bf C}:  The convergence correlations between individual pairs of
      replicates in \cite{Castro2019-uk} data, for the two selection lines
      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
      two regions Castro et al. (2019) found to have signals of parallel
      selection between LS1 and LS2. Through simulations, we have found
        that the differences in convergence correlation confidence interval
        widths between these \emph{Drosophila} studies and the Longshanks study
        are due to the differing population sizes.}

@@

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal
    covariances calculated in 100kb genomic windows from the
    Barghi et al. (2019) study, plotted alongside an empirical neutral null
    distribution created by recalculating the windowed covariances on 1,000
    sign permutations of allele frequency changes within tiles. The histogram
    bin number is 88, chosen by cross validation (SI Appendix, section
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
    windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
    $k=2 \times 10$ generations and in subfigure {\bf A} the covariances are
    separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal from
    the variance diagonal of the temporal covariance matrix (see cartoon of
    upper-triangle of covariance matrix in subfigures {\bf A} and {\bf B},
    where the first diagonal is the variance, and the dark gray indicates which
    off-diagonal of the covariance matrix is plotted in the histograms). {\bf
    C}: The lower and upper tail probabilities of the observed windowed
    covariances, at 20\% and 80\% quintiles of the empirical neutral null
    distribution, for varying time between allele frequency changes (i.e. which
    off-diagonal $k$). The confidence intervals are  $95\%$ block-bootstrap
  confidence intervals, and the light gray dashed line indicates the 20\% tail
probability expected under the neutral null. Similar figures for different
values of $k$ are in SI Appendix, section
\ref{suppfig:barghi-empnull-tilecovs}. }

@@

We observed that in the longest study  we analyzed \cite{Barghi2019-qy},
some genome-wide temporal covariances become negative at future timepoints (see
the first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
were on average going up initially are later going down in frequency, i.e.
that the average direction of selection experienced by alleles has flipped.
This might reflect either a change in the environment or the genetic
background, due to epistatic relationships among alleles altered by frequency
changes (which can occur during an optima shift; \cite{Hayward2019-kt}) or
recombination breaking up selective alleles. Such reversals in selection
dynamics could be occurring at other timepoints but the signal of a change in
the direction of selection at particular loci may be washed out when we
calculate our genome-wide average temporal covariances.  To address this
limitation, we calculated the distribution of the temporal covariances over
100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
pooling across all replicates; see SI Appendix, Fig.
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance estimate of each genomic window will be noisy, due to sampling and
genetic drift, and the neutral distribution of the covariance is complicated
due to linkage disequilibria, which can occur over long physical distances in
E\&R and selection studies \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
address this, we have developed a permutation-based procedure that constructs
an empirical neutral null distribution by randomly flipping the sign of
the allele frequency changes in each genomic window (i.e. a single random sign
flip is applied to all loci in a window). This destroys the systematic
covariances created by linked selection and creates a sampling distribution of
the covariances spuriously created by neutral genetic drift while preserving
the complex dependencies between adjacent loci created by linkage
disequilibrium.  This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone, as selection not only creates covariance between time
intervals, but also inflates the magnitude of allele frequency change within a
time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
empirical excess of windows with positive covariances between close timepoints
compared to the null distribution (a heavier right tail), and that this then
shifts to an excess of windows with negative covariances between more distant
timepoints (a heavier left tail)

@@

These simulation results are often relatively insensitive to the number of loci
underlying the trait, suggesting that they are capturing highly polygenic
signals.  Indeed if only a small number of loci influence the trait, the $G(t)$
trajectories are typically much more stochastic across replicates (compare
Figure \ref{fig:figure-1}B to SI Appendix, Fig.
\ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
selection response we see in the Barghi et al. (2019) data is highly
polygenic.  Using our simulations we find that sampling only every 10
generations does indeed mean that our estimates of $G(t)$ are an underestimate
of the contribution of linked selection as they cannot include the covariance
between closely spaced generations (see SI Appendix, Fig.
\ref{suppfig:supp-blocks}).

>>>>

\newcommand{\vb}[1]{{\it \color{red} #1}}

@@

\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
roles of genetic drift and selection in shaping genome-wide allele frequency
changes.  Selection can affect allele frequencies, both directly and
indirectly, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome, e.g. linked selection
\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
This approach has shown linked selection has a major role in shaping patterns
of genome-wide diversity across the genomes of a range of sexual species
\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over short
time intervals and \vb{of} its full impact on genome-wide allele frequency changes.

There are numerous examples of rapid phenotypic adaptation
\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature of
fitness makes detecting the impact of selection on genome-wide variation over
short timescales in sexual populations remarkably difficult \cite{Latta1998-me,
Pritchard2010-tk,Kemper2014-bx}. This is because the effect of selection on a
polygenic trait (such as fitness) is distributed across numerous loci. This can
lead to subtle allele frequency shifts on standing variation that are difficult
to distinguish from background levels of genetic drift and sampling variance.
Increasingly, genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large-effect selected loci \cite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous
paper, we proposed that one useful signal for understanding the genome-wide
impact of polygenic linked selection detectable from temporal genomic data is
the temporal autocovariance (i.e. covariance between two timepoints) \vb{of}
allele frequency changes \cite{Buffalo2019-io}. These covariances are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are expected to be zero for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a time interval
into the variances and covariances in allele frequency change between
generations.  Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an artificial
selection experiment \cite{Castro2019-uk}. These sequencing selection
experiments have started to uncover selected loci contributing to the adaptive
response; however it is as yet far from clear how much of \vb{genome-wide}
allele frequency \vb{changes} are driven by selection or genetic drift. We
repeatedly find a signal of temporal covariance, consistent with linked
selection acting to significantly perturb genome-wide allele frequency changes
across the genome in a manner that other approaches would not be able
differentiate from genetic drift. We estimate a lower bound of the fraction of
variance in allele frequency change caused by selection, as well as the
correlation between allele frequency changes between replicate populations
caused by convergent selection pressures. Overall, we demonstrate that linked
selection has a powerful role in shaping genome-wide allele frequency changes
over very short timescales \vb{in experimental evolution}.

@@

We first analyzed \vb{the dataset of} Barghi et al. (2019), an
evolve-and-resequence study with ten replicate populations exposed to a high
temperature lab environment and evolved for 60 generations, and sequenced every
ten generations. Using the seven timepoints and ten replicate populations, we
estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
for each of the ten replicates. Each row of these matrices represent the
temporal covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the
allele frequency change (in ten-generation intervals, denoted $\Delta_{_{10}}
p_t$) of some initial reference generation $s$ (the row of the matrix), and
some later timepoint $t$ (the column of the matrix). We corrected these
matrices for biases created due to sampling noise, and normalized the entries
for heterozygosity (see SI Appendix, sections \ref{supp:ind-depth-var-corr} and
\ref{supp:matrix-correction}). These covariances are expected to be zero when
only drift is acting, as only heritable variation for fitness can create
covariance between allele frequency changes in a closed population
\cite{Buffalo2019-io}.  Averaging across the ten replicate temporal covariances
matrices, we find temporal covariances that are statistically significant (95\%
block bootstraps CIs do not contain zero), consistent with linked selection
perturbing genome-wide allele frequency changes over very short time periods.
The covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals (Figure
\ref{fig:figure-1} A), as expected when directional selection affects linked
variants' frequency trajectories until ultimately linkage disequilibrium and
the associated additive genetic variance for fitness decays (which could occur
as a population reaches a new optimum, \vb{and directional selection weakens})
\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
this general pattern holds; see SI Appendix, Fig.
\ref{suppfig:barghi-cov-panels}.

Since our covariances are averages over loci, the covariance estimate could be
strongly affected by a few outlier regions. To test whether large outlier
regions drive the genome-wide signal we see in the Barghi et al. (2019) data,
we calculate the covariances in 100kb windows along the genome (we refer to
these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates (SI
Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
covariance, \vb{establishing} that genomic temporal covariances are non-zero
due to the impact of selection acting across many genomic regions.

@@

  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
    populations, through time from the Barghi et al. (2019) study. Each line
    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
    reference generation $s$ to a later time $t$ which varies along the x-axis;
    each line corresponds to a row of the upper-triangle of the temporal
    covariance matrix with the same color (upper right). The ranges around each
    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: \vb{A lower
    bound on the} proportion of the total variance in allele frequency change
    explained by linked selection, $G(t)$, as it varies through time $t$ along
    the x-axis. The black line is the $G(t)$ averaged across replicates, with
    the $95\%$ block-bootstrap confidence interval. The other lines are the
    $G(t)$ for each individual replicate, with colors indicating what subset of
    the temporal-covariance matrix to the right is being included in the
    calculation of $G(t)$.}

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change: the
ratio of total covariance in allele frequency change to the total variance in
allele frequency change. We denote the change in allele frequency as $\Delta
p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
Since the total variation in allele frequency change can be partitioned into
variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
\Delta p_j)$ (we \vb{correct for biases due to} sequencing depth), and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}.
\end{align}
%
\vb{This estimates the impact of selection on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time}. When the sum of the covariances
is positive, this measure can intuitively be understood as \vb{a lower bound on
} relative fraction of allele frequency change normally thought of as ``drift''
that is actually due to selection. Additionally, $G(t)$ can be understood as a
short-timescale estimate of the reduction in neutral diversity due to linked
selection (or equivalently the reduction in neutral effective population size
needed to account linked selection, see SI Appendix, section
\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
sequenced every ten generations, the numerator uses the covariances estimated
between ten-generation blocks of allele frequency change; thus the strong,
unobservable, covariances between adjacent generations do not contribute to the
numerator of $G(t)$. Had these covariances been measurable on shorter
timescales, their cumulative effect would likely have been higher yet (see SI
Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
details).  \vb{Additionally, selection inflates the variance in allele
frequency change per generation}; however, this effect cannot be easily
distinguished from drift. For both these reasons, our measure $G(t)$ is quite
conservative (we demonstrate this through simulations in SI Appendix, section
\ref{supp:tempblocks}).  Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection (Figure \ref{fig:figure-1} B). This proportion of
variance attributable to selection builds over time in Figure
\ref{fig:figure-1}B as the effects of linked selection are compounded over the
generations unlike genetic drift. Our G(t) starts to plateau to a constant
level as the covariances from earlier generations have decayed and so no longer
contribute as strongly (Figure \ref{fig:figure-1}).

@@

The replicate design of Barghi et al. (2019) allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are expected to
have allele frequency changes in the same direction if the selection pressures
are similar. Intuitively, where temporal covariances reflect that alleles
associated with heritable fitness backgrounds are predictive of frequency
changes between generations, replicate covariances reflect that heritable
fitness backgrounds common to each replicate predict (under the same selection
pressures) frequency changes between replicates; \vb{we note that there is not
  a direct one-to-one correspondence between temporal and replicate
  covariances, since the latter are driven by a shared selection pressure and
  the stochastic genetic backgrounds across replicate populations}. We measure
  this through a statistic similar to a correlation, which we call the
  convergent correlation: the ratio of average between-replicate covariance
  across all pairs to the average standard deviation across all pairs of
  replicates,

@@

We have calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2}A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2}A). We find these convergent correlation coefficients
are relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while a
substantial fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result consistent with the primary finding
of the original Barghi et al. (2019) study: that alternative loci
contribute to longer term adaptation across the different replicates.

@@

  \caption{{\bf A}: The convergence correlations, averaged across
    Barghi et al. (2019) replicate pairs, through time. Each line
    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
    p_{t})$ from a starting reference generation $s$ to a later time $t$, which
    varies along the x-axis; each line corresponds to a row of the temporal
    convergence correlation matrix depicted to the right (where the
      diagonal elements represent the convergence correlations between the same
      timepoints across replicate populations). We note that convergent
      correlation for the last timepoint is an outlier; we are unsure as to the
      cause of this, e.g. it does not appear to be driven by a single pair of
      replicates. {\bf B}: The convergence correlations between individual
      pairs of replicates in the Kelly and Hughes (2019) data (note the
      confidence intervals are plotted, but are small on this y-axis scale).
      {\bf C}:  The convergence correlations between individual pairs of
      replicates in \cite{Castro2019-uk} data, for the two selection lines
      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
      two regions Castro et al. (2019) found to have signals of parallel
      selection between LS1 and LS2. Through simulations, we have found
        that the differences in convergence correlation confidence interval
        widths between these \emph{Drosophila} studies and the Longshanks study
        are due to the differing population sizes.}

@@

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal
    covariances calculated in 100kb genomic windows from the
    Barghi et al. (2019) study, plotted alongside an empirical neutral null
    distribution created by recalculating the windowed covariances on 1,000
    sign permutations of allele frequency changes within tiles. The number of histogram
    bins is 88, chosen by cross validation (SI Appendix, section
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
    windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
    $k=2 \times 10$ generations and in subfigure {\bf A} the covariances are
    separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal from
    the variance diagonal of the temporal covariance matrix (see cartoon of
    upper-triangle of covariance matrix in subfigures {\bf A} and {\bf B},
    where the first diagonal is the variance, and the dark gray indicates which
    off-diagonal of the covariance matrix is plotted in the histograms). {\bf
    C}: The lower and upper tail probabilities of the observed windowed
    covariances, at 20\% and 80\% quintiles of the empirical neutral null
    distribution, for varying time between allele frequency changes (i.e. which
    off-diagonal $k$). The confidence intervals are  $95\%$ block-bootstrap
  confidence intervals, and the light gray dashed line indicates the 20\% tail
probability expected under the neutral null. Similar figures for different
values of $k$ are in SI Appendix, section
\ref{suppfig:barghi-empnull-tilecovs}. }

@@

We observed that in the longest study  we analyzed \cite{Barghi2019-qy},
some genome-wide temporal covariances become negative at future timepoints (see
the first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
were on average going up initially are later going down in frequency, i.e.
that the average direction of selection experienced by alleles has flipped.
This might reflect either a change in the environment or the genetic
background, due to epistatic relationships among alleles altered by frequency
changes (which can occur during an optima shift; \cite{Hayward2019-kt}) or
recombination breaking up selective alleles. Such reversals in selection
dynamics could be occurring at other timepoints but the signal of a change in
the direction of selection at particular loci may be washed out when we
calculate our genome-wide average temporal covariances.  To address this
limitation, we calculated the distribution of the temporal covariances over
100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
pooling across all replicates; see SI Appendix, Fig.
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance estimate of each genomic window will be noisy, due to sampling and
genetic drift, and the neutral distribution of the covariance is complicated
due to linkage disequilibria, which can occur over long physical distances in
E\&R and selection studies \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
address this, we have developed a permutation-based procedure that constructs
an empirical neutral null distribution by randomly flipping the sign of
the allele frequency changes in each genomic window (i.e. a single random sign
flip is applied to all loci in a window). This destroys the systematic
covariances created by linked selection and creates a sampling distribution of
the covariances spuriously created by neutral genetic drift while preserving
the complex dependencies between adjacent loci created by linkage
disequilibrium.  This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone, as selection not only creates covariance between time
intervals, but also inflates the magnitude of allele frequency change within a
time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
empirical excess of windows with positive covariances between close timepoints
compared to the null distribution (a heavier right tail), and that this then
shifts to an excess of windows with negative covariances between more distant
timepoints (a heavier left tail).

@@

\vb{Averaging across replicates, these simulation results show $G(t)$ is
  relatively insensitive to the number of loci underlying the trait. However,
  if only a small number of loci influence the trait, the $G(t)$ trajectories
  are typically much more stochastic \emph{across} replicates. This reaffirms
that the genome-wide linked selection response we see in the Barghi et al.
(2019) data is highly polygenic (compare Figure \ref{fig:figure-1}B to SI
Appendix, Fig.  \ref{suppfig:sim-expfit-Gs})}. \vb{Furthermore,} using our
simulations we find that sampling only every 10 generations does indeed mean
that our estimates of $G(t)$ are an underestimate of the \vb{proportional
effect} of linked selection as they cannot include the covariance
between closely spaced generations (see SI Appendix, Fig.
\ref{suppfig:supp-blocks}).
